Applying Radial Basis Networks and Markov Chains for on-line detection of concept drift in non-stationary environments

### 2019-10-27 tex/article/article.tex
Abstract and intro - Grammarly and translator checked

\begin{abstract}
Most real-world problems experience a phenomenon known as concept drift,
which is a change in data distribution that can affect the system performance.
%
However, the majority of drift detection methods are unsuited for non-stationary environments with data streams.
These algorithms or require the correct labeling of data - infeasible in these settings,
or do not match the severe response time and resource usage restrictions inherent to data streams.
%
In an attempt to mitigate the aforementioned problem,
this paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed method relies on Radial Basis Networks implicit clustering property
and uses Markov Chains to model the drifts transitions.
%
To assess the proposed method as a viable concept drift detector,
an analysis of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.
%
Furthermore, the algorithm was applied to the real-world problem of eye-tracking.
A relevant problem with impact in different areas of knowledge, since many behavioral experiments use eye-tracking information (fixations and saccades) as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time, with high accuracy, noise tolerance, and using limited resources.
\end{abstract}

@@

Still, most drift detection methods are improper for non-stationary environments with data streams.
These methods or require the correct labeling of data (impracticable in these contexts) and detect drift through the monitoring of the system performance,
or do not meet the severe response time and resource usage restrictions intrinsic to data streams.

This paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed algorithm is based on Radial Basis Networks implicit clustering property,
and employs Markov Chains to model the transitions and accomplish noise tolerance.

To validate the proposed method as a viable concept drift detector,
an examination of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.

Moreover, the algorithm was also tested with the real-world problem of eye-tracking.
This problem is relevant because many behavioral experiments use eye-tracking information (fixations and saccades) as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time, with high accuracy, noise tolerance, and using limited resources.

>>>>

\begin{abstract}
Most real-world problems experience a phenomenon known as concept drift,
a change in data distribution that can affect the system performance.
%
However, the majority of drift detection methods are unsuited for non-stationary environments with data streams.
These algorithms or require the correct labeling of data - infeasible in these settings,
or do not match the severe response time and resource usage restrictions inherent to data streams.
%
In an attempt to mitigate the aforementioned problem,
this paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed method relies on Radial Basis Networks implicit clustering property
and uses Markov Chains to model the drifts transitions.
%
To assess the proposed method as a viable concept drift detector,
an analysis of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.
%
Furthermore, the algorithm was applied to the real-world problem of eye-tracking.
A problem with impact in different areas of knowledge,
since many behavioral experiments use eye-tracking information as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time, with high accuracy, noise tolerance, and using limited resources.
\end{abstract}

@@

Still, most drift detection methods are improper for non-stationary environments with data streams.
These methods or require the correct labeling of data - impracticable in these contexts,
or do not meet the severe response time and resource usage restrictions intrinsic to data streams.

This paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed algorithm is based on Radial Basis Networks implicit clustering property
and employs Markov Chains to model the drifts transitions.
To validate the proposed method as a viable concept drift detector,
an examination of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.

Moreover, the algorithm was also applied to the real-world problem of eye-tracking.
A problem with impact in different areas of knowledge, since many behavioral experiments use eye-tracking information as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time, with high accuracy, noise tolerance, and using limited resources.
